---
title: "Visualization Exercise"
author: "Marcela Alfaro Córdoba"
format: html
editor: visual
---

## Let's try to make a plot pretty

Sometimes, the default settings of ggplot2 are not going to be enough for our taste, or for the requirements of the journal where we are publishing our results. This is an example of how to change the settings of a bubble chart.

## Let's start with the basic plot

We are going to use the data set gapminder, available in the package with the same name. Using ggplot2, we are going to construct a simple 3D chart using life expectancy, GDP per capita and population of each country.

```{r}
library(ggplot2)
library(dplyr)
library(gapminder)

data <- gapminder %>% filter(year == "2007") %>%  dplyr::select(-year)

ggplot(data, aes(x=gdpPercap, y = lifeExp, size = pop))+
  geom_point(alpha=0.7)
```

The first thing we need to change is how the population scale is represented. The scale is not very useful, since we cannot easily compare the numbers in the scientific format.

```{r}
data %>% 
  arrange(desc(pop)) %>% 
  mutate(country = factor(country, country)) %>% 
  ggplot(aes(x=gdpPercap, y = lifeExp, size = pop))+
  geom_point(alpha = 0.5) +
  scale_size(range = c(-.1, 24), name = "Population (M)")
```

But this didn't solve our original problem with the scale in the z axis. Now let's use color and try to modify the label.

```{r}
# Most basic bubble plot
data %>%
  arrange(desc(pop)) %>%
  mutate(country = factor(country, country)) %>%
  ggplot(aes(x=gdpPercap, y=lifeExp, size=pop, color=continent)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 24), name="Population (M)")
```

```{r}
data %>%
  arrange(desc(pop)) %>%
  mutate(country = factor(country, country)) %>%
  ggplot(aes(x=gdpPercap, y=lifeExp, size=pop, fill=continent)) +
    geom_point(alpha=0.5, shape=21, color="black") +
    geom_text(aes(x = gdpPercap, y = lifeExp + 3, label = country), # add some text annotations for the very large countries
              color = "grey50", size = 6,
              data = filter(data,pop > 1000000000 | 
                              country %in% c("Nigeria", "United States"))) +
  scale_x_log10(limits = c(200, 60000)) + # clean the axes names and breaks
    scale_size(range = c(.1, 24), name="Population (M)", guide = "none") +
    scale_fill_viridis_d() +
    theme_linedraw() +
    ylab("Life Expectancy") +
    xlab("Log Gdp per Capita") +
  ggtitle("Life Expectancy, GDP per capita and population, 2007") +
  theme(
   legend.position = "bottom",
  plot.title = element_text(color = "black", size = 12, face = "bold"),
  text = element_text(family = "mono"))
```

You can check a list of fonts in here: http://www.cookbook-r.com/Graphs/Fonts/
