---
title: "Quarto Workshop"
author: Marcela Alfaro Córdoba
format: html
editor: visual
---

## Quarto

This is our first report using Quarto. We are going to read a data set that we have in our project folder, we are going to explore it numerically and graphically, and then we are going to change the data to see how we can easily get it updated.

## Installing packages and reading data

We are going to use SAFI.csv, because that's what we have available. However, you can follow this code to read any data set that you have in your project files.

```{r pack_data, include=FALSE}
library("tidyverse")
SAFI_clean <- read_csv("data/SAFI_clean.csv")
options(dplyr.summarise.inform = FALSE)
```

## EDA

This report uses the **tidyverse** package along with the *SAFI* dataset, which has columns that include:

-   village
-   interview_date
-   no_members
-   memb_assoc
-   years_liv
-   respondent_wall_type
-   rooms

```{r echo=FALSE, include=FALSE, warning=FALSE}
SAFI <- SAFI_clean %>% select(village,interview_date,no_membrs,memb_assoc, years_liv,respondent_wall_type,rooms) 
str(SAFI)
```

Now, let's create a table that summarizes the mean household size by village

```{r echo=FALSE, warning=FALSE}
SAFI %>%
    filter(!is.na(memb_assoc)) %>%
    group_by(village, memb_assoc) %>%
    dplyr::summarize(mean_no_membrs = mean(no_membrs)) %>%
  knitr::kable(caption = "SAFI: Mean number of members per association and village", col.names = c("Village", "Member Association", 
                             "Mean Number of Members"))

# create a summary data frame with the mean household size by village
mean_household <- SAFI_clean %>%
    group_by(village) %>%
    dplyr::summarize(mean_no_membrs = mean(no_membrs))

# and select the village we want to use
mean_chirodzo <- mean_household %>%
  filter(village == "Chirodzo")
```

The average household size in the village of Chirodzo is `r round(mean_chirodzo$mean_no_membrs, 2)`

## Plots

```{r include=FALSE}
interviews_plotting <- SAFI_clean %>%
  ## pivot wider by items_owned
  separate_rows(items_owned, sep = ";") %>%
  ## if there were no items listed, changing NA to no_listed_items
  replace_na(list(items_owned = "no_listed_items")) %>%
  mutate(items_owned_logical = TRUE) %>%
  pivot_wider(names_from = items_owned, 
              values_from = items_owned_logical, 
              values_fill = list(items_owned_logical = FALSE)) %>%
  ## pivot wider by months_lack_food
  separate_rows(months_lack_food, sep = ";") %>%
  mutate(months_lack_food_logical = TRUE) %>%
  pivot_wider(names_from = months_lack_food, 
              values_from = months_lack_food_logical, 
              values_fill = list(months_lack_food_logical = FALSE)) %>%
  ## add some summary columns
  mutate(number_months_lack_food = rowSums(select(., Jan:May))) %>%
  mutate(number_items = rowSums(select(., bicycle:car)))
```

```{r chunk-name, echo=FALSE, fig.cap = "I made this plot while attending an awesome Data Carpentries workshop where I learned a ton of cool stuff!"}
interviews_plotting %>%
  ggplot(aes(x = respondent_wall_type)) +
  geom_bar(aes(fill = village), position = "dodge") + 
  labs(x = "Type of Wall in Home", y = "Count", fill = "Village Name") +
  scale_fill_viridis_d() # add colour deficient friendly palette
```

What happens if there's a correction in the data table? Let's do it:

```{r}
SAFI2 <- SAFI_clean[-c(85,86),]
write.csv(SAFI2, file="data/SAFI2.csv")
```

Now let's change SAFI_clean for SAFI2 in the doc..

Using python: https://quarto.org/docs/computations/python.html

More references: https://quarto.org/
