---
title: "My Awesome Report"
author: "Marcela Alfaro Córdoba"
date: ' '
output:
  html_document:
    df_print: paged
---
  
```{r setup, include = FALSE}
library(tidyverse)
library(here)
interviews <- read_csv(here("data/SAFI_clean.csv"), na = "NULL")
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This report uses the **tidyverse** package along with the *SAFI* dataset, which has columns that include:

- village
- interview_date
- no_members
- years_liv
- respondent_wall_type
- rooms

Or if you prefer:

1. village
2. interview_date
3. no_members
4. years_liv
5. respondent_wall_type
6. rooms

Or if you want to add sections:

- village
  + Name of village
- interview_date
  + Date of interview
- no_members
  + How many family members lived in a house
- years_liv
  + How many years respondent has lived in village or neighbouring village
- respondent_wall_type
  + Type of wall of house
- rooms
  + Number of rooms in house


```{r warning = FALSE, message = FALSE}
interviews %>%
    filter(!is.na(memb_assoc)) %>%
    group_by(village, memb_assoc) %>%
    summarize(mean_no_membrs = mean(no_membrs)) %>%
  knitr::kable(caption = "We can also add a caption.", 
               col.names = c("Village", "Member Association", 
                             "Mean Number of Members"))
```

Now we can make an informative statement on the means of each village, and include the mean values as in-line R-code. For example:

```{r include=FALSE}
# create a summary data frame with the mean household size by village
mean_household <- interviews %>%
    group_by(village) %>%
    summarize(mean_no_membrs = mean(no_membrs))

# and select the village we want to use
mean_chirodzo <- mean_household %>%
  filter(village == "Chirodzo")
```


The average household size in the village of Chirodzo is `r round(mean_chirodzo$mean_no_membrs, 2)`


```{r include=FALSE}
## Not run, but can be used to load in data from previous lesson!
interviews_plotting <- interviews %>%
  ## pivot wider by items_owned
  separate_rows(items_owned, sep = ";") %>%
  ## if there were no items listed, changing NA to no_listed_items
  replace_na(list(items_owned = "no_listed_items")) %>%
  mutate(items_owned_logical = TRUE) %>%
  pivot_wider(names_from = items_owned, 
              values_from = items_owned_logical, 
              values_fill = list(items_owned_logical = FALSE)) %>%
  ## pivot wider by months_lack_food
  separate_rows(months_lack_food, sep = ";") %>%
  mutate(months_lack_food_logical = TRUE) %>%
  pivot_wider(names_from = months_lack_food, 
              values_from = months_lack_food_logical, 
              values_fill = list(months_lack_food_logical = FALSE)) %>%
  ## add some summary columns
  mutate(number_months_lack_food = rowSums(select(., Jan:May))) %>%
  mutate(number_items = rowSums(select(., bicycle:car)))
```


```{r chunk-name, echo=FALSE,fig.cap = "I made this plot while attending an awesome Data Carpentries workshop where I learned a ton of cool stuff!"}
interviews_plotting %>%
  ggplot(aes(x = respondent_wall_type)) +
  geom_bar(aes(fill = village), position = "dodge") + 
  labs(x = "Type of Wall in Home", y = "Count", fill = "Village Name") +
  scale_fill_viridis_d() # add colour deficient friendly palette
```


